---
title: "goatscape_pkg_notes"
author: "Bryan Milstead"
date: "`r format(Sys.time(), '%d %B %Y')`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## to do
* is [goatscape_pkg_notes.rmd] in the correct directory? [inst/extdata/b]?

## Introduction

This .rmd file documents the work process for building the package [goatscape]

## installation
library("devtools")
install_github("usepa/goatscape")
library("goatscape")

## get_census

### Data Files
* r-script used to create the package data [data-raw/create_pkg_data.r]

### geoid is the unique identifier for census blocks/block ids; not currently using it and varies by year
* The geoid is a combination of "state fips", "county fips", "tract id", and "block id".  For example: geoid == 440030208003077 corresponds to statefp == 44, countfp == 003, tractce == 020800, and blockce == 3077.

* Note: there can be differences among census years.

### census codes
* Total Population
    - name == "sf3"; vintage == 2000; vars == "P001001"
    - name == "sf1"; vintage == 2010; vars == "P0010001"
    
## get_nlcd
* NLCD website: http://www.mrlc.gov/index.php
* basic example: http://spatialecology.weebly.com/r-code--data/category/data-vegetation
* The National Map API: https://viewer.nationalmap.gov/help/documents/TNMAccessAPIDocumentation/TNMAccessAPIDocumentation.pdf
* feddata package being updated; good code to copy: https://github.com/ropensci/FedData/tree/master/R
* NLCD data on AWS; tiled by bbox: https://s3-us-west-2.amazonaws.com/prd-tnm/StagedProducts/NLCD/data/2011/landcover/3x3/NLCD2011_LC_N24W078.zip
* NLCD USGS: https://www.sciencebase.gov/catalog/item/5a1c31abe4b09fc93dd6381c

### message from Kyle Walker about [tigris]

Kyle Walker <notifications@github.com>
5:52 PM (17 hours ago)

to walkerke/tigris, me, Author 
I'm glad you've found the package useful! This is a limitation of how the data are provided by the Census Bureau. Block shapefiles are available by state; tigris downloads the state shapefile then subsets by county in R. As such there isn't a way to speed up the download outside of a faster internet connection. However, I'd recommend:

Using caching with options(tigris_use_cache = TRUE);
Using the simple features class with class = "sf" or options(tigris_class = "sf"); this speeds loading times signficantly (especially when paired with shapefile caching);
Identifying the blocks - or region of potential blocks - you need and saving out as a .rds file, then loading back in as needed.


```{r example}
library(raster)

#Download data
temp <- tempfile() #create a temp file

#download the data (this is a 1 TB file)
download.file("http://gisdata.usgs.gov/TDDS/DownloadFile.php?TYPE=nlcd2011&FNAME=nlcd_2011_landcover_2011_edition_2014_10_10.zip",temp)

#unzip the file
unzip(temp, exdir=here::here('temp'))

#delete the temp file
unlink(temp)

#Rasterize the data (package “raster”)
  cont_usa <- raster(here::here("temp/nlcd_2011_landcover_2011_edition_2014_10_10/nlcd_2011_landcover_2011_edition_2014_10_10.img"))

#Change the resolution of the data to decrease file size using aggregate()
  cont_usa_coarse <- aggregate(Cont_USA, fact=100, fun=max)

#Plot the NLCD data
  plot(cont_usa_coarse, main="NLCD Land Cover Types; Continental USA; full res", legend = FALSE, asp=1)
  plot(cont_usa_coarse, main="NLCD Land Cover Types; Continental USA; lower res", legend = FALSE, asp=1)
  
```




